\begin {document}
\section{Generative Models for Peptide Design}

To generate novel peptide candidates, two generative architectures were implemented: a Variational Autoencoder (VAE) and a Conditional Variational Autoencoder (CVAE). Both models operate on peptide sequences encoded as fixed-length one-hot vectors, where each sequence position is represented over the 20 canonical amino acids. For a peptide of length $L$, the input dimensionality is therefore $20 \times L$.

The objective of these models is to learn a structured latent representation of peptide sequences that can be sampled in order to generate new candidates. The VAE provides unconstrained generation based on the global training distribution, whereas the CVAE extends this framework by conditioning generation on physicochemical descriptors, enabling more directed design.

\subsection{Variational Autoencoder}

The VAE is a probabilistic generative model that learns to encode peptide sequences into a continuous latent space and decode latent samples back into amino acid sequences. Given an input peptide representation $x$, the encoder predicts the parameters of a Gaussian latent distribution:
\begin{equation}
q_{\phi}(z \mid x) = \mathcal{N}(\mu_{\phi}(x), \sigma_{\phi}^{2}(x)).
\end{equation}

A latent vector $z$ is then sampled using the reparameterization trick:
\begin{equation}
z = \mu + \sigma \odot \epsilon, \qquad \epsilon \sim \mathcal{N}(0, I),
\end{equation}
which allows gradients to propagate through the sampling step during training.

In the implemented architecture, the encoder is a multilayer perceptron composed of linear layers, layer normalization, and ReLU activations. The output is split into two vectors corresponding to the latent mean and log-variance. The decoder is another multilayer perceptron of symmetric structure that maps the latent vector back to the original sequence space.

Rather than reconstructing the sequence using a mean squared error loss, the decoder outputs raw logits for each amino acid at each sequence position. These logits are reshaped into categorical distributions over the 20 amino acids, and the reconstruction loss is computed using cross-entropy. This choice is more appropriate for peptide sequences, since amino acids are discrete categories rather than continuous values.

The training objective combines reconstruction quality and latent regularization:
\begin{equation}
\mathcal{L}(x) = \mathcal{L}_{\text{recon}}(x, \hat{x}) + \beta \, D_{\text{KL}}\bigl(q_{\phi}(z \mid x)\, \|\, p(z)\bigr),
\end{equation}
where $p(z)=\mathcal{N}(0,I)$ is the prior distribution and $D_{\text{KL}}$ denotes the Kullback--Leibler divergence. In practice, the KL term is progressively increased during training through KL annealing in order to stabilize optimization and reduce posterior collapse at early epochs.

The model is optimized with Adam, combined with cosine annealing of the learning rate and gradient clipping to improve training stability. During generation, latent vectors are sampled from the standard normal distribution and decoded into amino acid probability distributions. Sequence diversity is controlled through a temperature parameter applied before softmax sampling.

\subsection{Conditional Variational Autoencoder}

The CVAE extends the VAE by conditioning both encoding and decoding on an auxiliary vector $y$ describing target peptide properties. Instead of modeling only the unconditional peptide distribution, the CVAE learns a conditional distribution:
\begin{equation}
q_{\phi}(z \mid x, y), \qquad p_{\theta}(x \mid z, y).
\end{equation}

In the implemented model, the condition vector is concatenated with the peptide input before the encoder and concatenated again with the latent representation before the decoder. This architecture allows the model to generate peptides not only consistent with the training distribution, but also biased toward a specified physicochemical profile.

The condition vector is constructed from a fixed ordered list of physicochemical descriptors:
\begin{itemize}
\item length,
\item molecular weight,
\item net charge at pH 5.5,
\item isoelectric point,
\item hydrophobicity,
\item hydrophobic moment,
\item logP,
\item Boman index,
\item number of hydrogen bond donors,
\item number of hydrogen bond acceptors,
\item topological polar surface area (TPSA).
\end{itemize}

Each property is encoded using two values corresponding to a minimum and maximum target:
\begin{equation}
[\text{min}, \text{max}].
\end{equation}
If only a scalar value is provided, the same value is used for both entries. Missing values are set to zero. The resulting vector is either padded with zeros or truncated in order to match the predefined conditioning dimensionality.

This conditioning strategy enables the user to steer generation toward peptides with desired physicochemical characteristics. For example, the model can be guided toward sequences with a target range of charge, hydrophobicity, or molecular weight. In that sense, the CVAE is better suited than the VAE for directed peptide design.

\subsection{Comparison Between VAE and CVAE}

The VAE and CVAE serve related but distinct purposes. The VAE is an unconditional model: it learns the overall distribution of the training peptides and can sample novel sequences from this learned space. It is therefore useful for exploratory generation and for assessing whether the latent space captures relevant structural patterns in peptide sequences.

By contrast, the CVAE introduces control over the generation process. Because the model is explicitly conditioned on target descriptors, it can be used to bias sampling toward peptides exhibiting specific physicochemical properties. This makes the CVAE more appropriate for inverse design settings, where the goal is not merely to generate plausible peptides, but to generate candidates with a desired profile.

In summary, the VAE emphasizes diversity and distribution learning, whereas the CVAE emphasizes controllability and property-guided generation.

\subsection{Physicochemical Conditioning}

The physicochemical descriptors used by the CVAE can be computed from peptide sequences using the chemistry analysis module. These descriptors include molecular weight, charge, isoelectric point, hydrophobicity, hydrophobic moment, Boman index, and molecular descriptors derived from structure such as logP, hydrogen bond donor and acceptor counts, and TPSA.

This link between the chemistry module and the generative model is important, because it provides a principled way to translate biochemical objectives into numerical conditioning variables. As a result, the CVAE can be integrated into a broader peptide design pipeline in which candidate peptides are generated, analyzed, and filtered according to downstream criteria.

\subsection{Limitations}

Several limitations must be acknowledged. First, the current implementation assumes fixed-length peptide inputs, since the one-hot encoding and decoder output dimensionality depend directly on sequence length. As a result, the model does not naturally support variable-length peptides without additional preprocessing or architectural changes.

Second, the conditioning mechanism in the CVAE should be interpreted as a soft constraint rather than a hard constraint. The condition vector influences the probability distribution learned by the decoder, but it does not guarantee that every generated peptide will strictly satisfy the requested bounds. Consequently, an additional post-generation filtering step may still be necessary to enforce physicochemical thresholds.

Third, the usefulness of the CVAE depends strongly on the quality and consistency of the condition labels used during training. If the model is trained on sequences that are not paired with meaningful physicochemical targets, the conditioning mechanism will not learn a reliable relationship between conditions and generated sequences.

Finally, neither the VAE nor the CVAE directly predicts antimicrobial efficacy or biological activity. These models generate plausible peptide candidates from a learned sequence distribution, but downstream validation remains necessary to assess activity, selectivity, toxicity, and synthetic feasibility.

\subsection{Conclusion}

Overall, the VAE and CVAE provide two complementary approaches for peptide generation. The VAE offers a flexible framework for learning a continuous latent representation of peptide sequences and exploring the learned sequence space. The CVAE extends this framework by incorporating physicochemical guidance, making it more suitable for directed generative design. Together, these models form a useful foundation for computational peptide discovery, particularly when combined with chemical analysis and post-generation filtering.
\end{document}