\section{Introduction}\label{sec:introduction}

Peptides, and especially antimicrobial peptides (AMPs), are promising candidates for therapeutic, agricultural, and industrial applications \cite{wang2022machine}. Their \textit{in silico} design is typically formulated as an iterative optimization process: generate candidate sequences, apply formal constraints, and evaluate each candidate with proxy scores for activity, toxicity risk, and stability \cite{wang2022machine,veltri2018deep}. While recent machine learning methods have accelerated sequence design, balancing multiple and sometimes conflicting objectives remains difficult in practice.

This project addresses that challenge through an agentic proof of concept in which specialized software agents cooperate in a closed optimization loop. The pipeline includes an AI Designer agent that proposes candidates, a Chemistry Constraints agent that filters feasibility using proxy physicochemical rules, and a Biology Function/Risk agent that critiques expected function and risk. An orchestrator manages the iterative cycle (propose $\rightarrow$ filter $\rightarrow$ score $\rightarrow$ revise), applies stopping criteria, and stores logs for traceability.

The work is centered on AI aspects (modeling, orchestration, reasoning, and quantitative evaluation) and assumes no prior expertise in wet-lab biology or chemistry. Evaluation is explicit and reproducible: we measure validity, average and Top-K score, diversity, pass rates for constraints, and score evolution across rounds, then compare against a non-agentic baseline. This framing allows us to test whether multi-expert coordination provides measurable gains over single-pass generation .

The remainder of this paper is structured as follows: Section \ref{sec:context} provides a brief overview of the project context and related work. Section \ref{sec:architecture} details the architecture of our multi-agent pipeline and the specific roles of the AI, Chemistry, and Biology agents. Section \ref{sec:setup} outlines our experimental setup and quantitative evaluation metrics. Section \ref{sec:results} presents our findings, including baseline comparisons and ablation studies. Finally, Section \ref{sec:conclusion} concludes the report and discusses potential future work.