\section{Experimental Setup}\label{sec:setup}
This section documents the implemented scoring and comparison protocol used in the current proof of concept.

\subsection{Quantitative Scoring System}\label{subsec:scoring}

\subsubsection{Biologist Agent Scoring}
The objective is to rank a batch of peptides and identify which candidate is closest to the provided context (an existing peptide or requested properties represented as a reference embedding).

The biologist module uses ESM-2 embeddings (default: \url{facebook/esm2\_t12\_35M\_UR50D}) with attention-mask-aware mean pooling to produce one vector per peptide.

Two proxy variants are implemented. The first uses cosine similarity, mapped to $[0,1]$. The second (Global L2) computes Euclidean distance to a reference embedding and applies a batch-independent exponential mapping. In preliminary tests, cosine scores showed limited contrast for short peptides, so the Global L2 variant was preferred for iterative ranking.

The resulting score is used by the orchestrator for candidate ranking and Top-$K$ selection. It is treated as a proxy signal to guide search, not as an absolute biological activity measurement.

\subsubsection{Chemist Agent Scoring}

To prevent pipeline starvation (where strict filters might discard the entire population early in the optimization process), the orchestrator applies a simple fallback policy: if no in-limits candidate is available after chemistry evaluation, biological scoring is computed on the full chemist output (including out-of-limits candidates).

The chemistry workflow therefore operates in two practical steps:
\begin{enumerate}
    \item \textbf{Hard constraint check and scoring:} Candidates are evaluated against strict limits and scored from target-distance-based property scores.
    \item \textbf{Fallback at orchestration level:} If no candidate satisfies all limits, out-of-limits candidates are still passed to biology scoring to keep the iterative loop active.
\end{enumerate}
This design ensures a continuous flow of candidates to the Biology Agent while maintaining selection pressure toward developable molecules.

\begin{itemize}
    \item \textbf{Solubility \& Stability proxies:} Isoelectric point and net charge are computed for each peptide.
    \item \textbf{Permeability proxy:} The profile includes $\log P$ (RDKit-based) and hydrophobicity.
    \item \textbf{Basic developability descriptors:} Length and molecular weight are included in scoring and admissibility checks.
\end{itemize}

\subsubsection{Orchestrator-Level Scoring and Ranking}
At orchestration level, each candidate receives a combined proxy score obtained by averaging Chemistry and Biology signals with equal weight:
\begin{equation}
S_{\text{combined}} = \frac{S_{\text{chem}} + S_{\text{bio}}}{2}.
\end{equation}
The ranking policy then prioritizes chemically admissible peptides (in-limits) and, within that constraint status, orders candidates by combined score (with expert scores used as tie-breakers). The orchestrator maintains this ranking globally across rounds, retaining one best entry per unique sequence and returning the final Top-$K$ from this global pool.

In practical terms, this means candidate quality is not decided only round by round: a strong peptide discovered early can remain competitive at the end if later rounds do not surpass it.

\subsection{Evaluation Metrics}\label{subsec:metrics}
We report six core metrics, computed at run level and on final Top-$K$ outputs:
\begin{itemize}
    \item \textbf{Validity rate (\%):} fraction of returned peptides that satisfy all configured chemistry constraints (\texttt{in\_limits=True}).
    \item \textbf{Top-1 score:} best combined score among returned candidates.
    \item \textbf{Top-$K$ mean score:} average combined score across final Top-$K$ peptides.
    \item \textbf{Constraint pass rate:} proportion of generated peptides passing chemistry constraints before final ranking.
    \item \textbf{Diversity (sequence-level):} mean pairwise distance (normalized edit distance) and unique-ratio in Top-$K$.
    \item \textbf{Diversity (distribution-level):} bigram entropy over amino-acid pairs as a complementary compositional diversity proxy.
\end{itemize}

At each iteration, chemistry and biology signals are merged through the orchestrator-level score:
\begin{equation}
S_{\text{combined}} = \frac{S_{\text{chem}} + S_{\text{bio}}}{2}.
\end{equation}
The final ranking is built globally across rounds (best-by-sequence retention), then truncated to Top-$K$.

\subsection{Baselines}\label{subsec:baselines}
The baseline is designed as a simpler non-agentic reference to test whether the proposed multi-expert iterative pipeline provides measurable gains. Concretely, we use a standalone Conditional Variational Autoencoder (CVAE) that directly generates peptide candidates from conditioning features, without chemistry/bio critique loops and without orchestrated revision rounds.
This baseline CVAE is trained on the dataset built with \path{database/get_data.py}. The extraction script queries DBAASP asynchronously and retains sequence-level and physicochemical fields used for conditioning (length, pH, molecular weight, logP, net charge, isoelectric point, hydrophobicity, and cationicity). 


\begin{figure}[H]
\centering
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[
    >=latex,
    modelBlock/.style={
        rectangle,
        thick,
        minimum width=3cm,
        minimum height=1.5cm,
        align=center,
        font=\sffamily\bfseries,
        rounded corners=2pt
    },
    dataBlock/.style={
        rectangle,
        thick, dashed,
        minimum width=2.5cm,
        minimum height=1cm,
        align=center,
        font=\sffamily\footnotesize,
        rounded corners=2pt,
        fill=gray!5, draw=gray!60
    },
    condBlock/.style={
        rectangle,
        thick,
        minimum width=4.5cm, % <--- Reduced from 6.6cm
        minimum height=1.2cm,
        align=center,
        font=\sffamily\footnotesize,
        rounded corners=2pt,
        fill=orange!10, draw=orange!80!black
    },
    encStyle/.style={modelBlock, fill=blue!10, draw=blue!80!black},
    decStyle/.style={modelBlock, fill=green!10, draw=green!80!black},
    latentStyle/.style={modelBlock, fill=purple!10, draw=purple!80!black, minimum width=2.7cm},
    outStyle/.style={dataBlock, solid, fill=yellow!15, draw=yellow!60!black},
    flowArrow/.style={-{Stealth[length=3mm]}, thick, draw=black!90},
    condArrow/.style={-{Stealth[length=3mm]}, thick, draw=orange!90!black},
    inferArrow/.style={-{Stealth[length=3mm]}, thick, dotted, draw=purple!80!black}
]

\node[dataBlock] (inputSeq) at (-6, 0) {Input Sequence $x$\\ \scriptsize(Training)};
\node[encStyle] (encoder) at (-3, 0) {Encoder\\ \scriptsize(Embedding + GRU)};
\node[latentStyle] (latent) at (0, 0) {Latent \\ \scriptsize$z=\mu+\sigma\odot\epsilon$};
\node[decStyle] (decoder) at (3, 0) {Decoder\\ \scriptsize(Autoregressive GRU)};
\node[outStyle] (outputSeq) at (6, 0) {Generated\\ peptide $\hat{x}$};

% Added a line break (\\) after "net charge,"
\node[condBlock] (condition) at (0, -2.5) {\textbf{Condition vector} $c$\\ \scriptsize [length, pH, MW, logP, net charge, \\ pI, hydrophobicity, cationicity]};

\node[dataBlock, draw=purple!60, fill=purple!5] (noise) at (0, 2.5) {Prior sample\\ $\mathcal{N}(0, I)$\\ \scriptsize(Inference)};

\draw[flowArrow] (inputSeq) -- (encoder);
\draw[flowArrow] (encoder) -- (latent);
\draw[flowArrow] (latent) -- (decoder);
\draw[flowArrow] (decoder) -- (outputSeq);

\draw[condArrow] (condition.west) -| node[pos=0.75, left, font=\sffamily\scriptsize, text=orange!80!black] {concat $c$} (encoder.south);
\draw[condArrow] (condition.east) -| node[pos=0.75, right, font=\sffamily\scriptsize, text=orange!80!black] {concat $c$} (decoder.south);

\draw[inferArrow] (noise) -- (latent);

% Symmetric bounding box to enforce true horizontal centering.
\pgfresetboundingbox
\path (-6.025, -3.3) rectangle (7.8, 2.8);

\end{tikzpicture}
}
\caption{Baseline CVAE architecture used for comparison. During training, the model learns $q_{\phi}(z\mid x,c)$ and reconstructs sequences with an autoregressive decoder conditioned on $z$ and $c$. During inference, peptides are generated by sampling $z\sim\mathcal{N}(0,I)$ and decoding once, without iterative multi-agent feedback.}
\label{fig:baseline_cvae}
\end{figure}


In practice, many API entries contain missing values for one or more properties. The baseline data pipeline therefore substitutes absent fields with fixed defaults before standardization. This dataset is thus not optimal from a data-quality perspective, but for a proof-of-concept objective we considered it sufficient to establish a consistent comparison point.
From an implementation perspective, the baseline model is a GRU-based CVAE. The encoder embeds amino-acid tokens and maps sequence-condition pairs to latent Gaussian parameters; the decoder is autoregressive and conditioned on both latent code and property vector. Training uses cross-entropy reconstruction loss with KL regularization, Adam optimization, and a progressive KL weight (annealed $\beta$).
At inference time, candidate peptides are sampled from the latent prior under a target condition vector (scaled with the training \texttt{StandardScaler}), with temperature and top-$k$ sampling to control exploration. This baseline serves only as the comparison anchor: unlike the full pipeline, it performs one-shot conditional generation and does not benefit from multi-agent feedback, constraint-aware iteration, or round-level refinement.
