\begin{thebibliography}{9}

\bibitem{wan2024machine}
F. Wan, F. Wong, J. J. Collins, and C. de la Fuente-Nunez, ``Machine learning for antimicrobial peptide identification and design,'' \textit{Nature Reviews Bioengineering}, Feb. 2024. [Online]. Available: https://doi.org/10.1038/s44222-024-00152-x

\bibitem{rives2021biological}
A. Rives, J. Meier, T. Sercu, S. Goyal, Z. Lin, J. Liu, D. Guo, M. Ott, C. L. Zitnick, J. Ma, and R. Fergus, ``Biological Structure and Function Emerge from Scaling Unsupervised Learning to 250 Million Protein Sequences,'' \textit{Proceedings of the National Academy of Sciences}, vol. 118, no. 15, p. e2016239118, 2021.

\bibitem{xi2023agentic}
Z. Xi, et al., ``The Rise and Potential of Large Language Model Based Agents: A Survey,'' \textit{arXiv preprint arXiv:2309.07864}, 2023.

\bibitem{guo2024multiagent}
T. Guo, et al., ``Large Language Model based Multi-Agent Systems: A Survey of Progress and Challenges,'' \textit{arXiv preprint arXiv:2401.03428}, 2024.

\bibitem{boiko2023coscientist}
D. A. Boiko, R. MacKnight, and G. Gomes, ``Autonomous chemical research with large language models,'' \textit{Nature}, vol. 624, pp. 570--578, 2023.

\bibitem{huang2025biomni}
K. Huang, et al., ``Biomni: A General-Purpose Biomedical AI Agent,'' \textit{bioRxiv}, 2025.

\bibitem{wang2022machine}
L. Wang \textit{et al.}, ``Machine learning for peptide design,'' \textit{Nature Computational Science}, vol.~2, pp.~100--112, 2022.

\bibitem{veltri2018deep}
D. Veltri, U. Kamath, and A. Shehu, ``Deep learning improves antimicrobial peptide recognition,'' \textit{Bioinformatics}, vol.~34, no.~16, pp.~2740--2747, 2018.

\bibitem{esmgithub}
Facebook Research,
``ESM: Evolutionary Scale Modeling,'' GitHub repository. [Online]. Available: \url{https://github.com/facebookresearch/esm}

\end{thebibliography}